---
title: "week2 assignment1"
author: "Jingwen Mou"
date: "2021/11/5"
output: html_document
---

###0. Load the data and set global options.

```{r, echo = TRUE, message = FALSE, warning=FALSE}
setwd("/Users/apple/Desktop/datasciencecoursera/notes/Reproducible Research/week2/a1")
whole_data <- read.csv("activity.csv")
library(dplyr)
library(ggplot2)
```


###1.What is mean total number of steps taken per day? 

```{r, echo = TRUE, warning = FALSE}
whole_data_clean <- na.omit(whole_data)
total_by_date <- aggregate(steps ~ date, whole_data_clean, sum)
hist(total_by_date$steps,
     main = "Total Number of Steps Taken Each Day",
     xlab = "Number of Steps Taken")
mean_steps <- sum(total_by_date$steps)/61
median_steps <- median(total_by_date$steps)
```

The mean total number of steps taken per day is `r mean_steps`.
The median total number of steps taken is `r median_steps`.


###2.What is the average daily activity pattern?

```{r, echo = TRUE, warning = FALSE}
daily_pattern <- aggregate(steps ~ interval, whole_data_clean, sum)
daily_pattern$steps <- daily_pattern$steps/61
with(daily_pattern, plot(interval, steps,
                         main = "Average Daily Activity Pattern",
                         xlab = "Time Interval",
                         ylab = "Average Steps Taken",
                         type = "l"))
max_steps_row <- which.max(daily_pattern$steps)
max_steps_interval <- daily_pattern[max_steps_row, ]$interval
max_steps <- daily_pattern[max_steps_row, ]$steps
```

The 5-minute interval containing the maximum number of steps is `max_steps_inerval`, when `max_steps` steps are taken in that interval.


###3. Imputing missing values.

```{r, echo = TRUE, warning = FALSE}
na_row_number <- nrow(whole_data) - nrow(whole_data_clean)
```

There are `na_row_number` rows with missing values.

```{r, echo = TRUE, warning = FALSE}
whole_data1 <- whole_data
na_position <- is.na(whole_data1$steps)
avg_steps <- tapply(whole_data1$steps, whole_data1$interval, mean,
                    na.rm =TRUE, simplify = TRUE)
whole_data1$steps[na_position] <- avg_steps[as.character
                                (whole_data1$interval[na_position])]
```

```{r, echo = TRUE, warning = FALSE}
total_by_date1 <- aggregate(steps ~ date, whole_data1, sum)
hist(total_by_date1$steps,
     main = "Total Number of Steps Taken Each Day #2",
     xlab = "Number of Steps Taken")
mean_steps1 <- sum(total_by_date1$steps)/61
median_steps1 <- median(total_by_date1$steps)
```

The mean total number of steps taken per day is `r mean_steps1`.
The median total number of steps taken is `r median_steps1`.
The results differ from the estimates from the first part, the mean and median are the same this time. The estimates using this startegy is more accurate.


###4. Are there differences in activity patterns between weekdays and weekends?


```{r, echo = TRUE, warning = FALSE}
whole_data1$date <- as.Date(whole_data1$date)

whole_data1 <- mutate(whole_data, Weekdays = ifelse(weekdays(whole_data1$date) == "星期六" | weekdays(whole_data1$date) == "星期日", "weekend", "weekday"))

whole_data1$Weekdays <- as.factor(whole_data1$Weekdays)
```

```{r, echo = TRUE, warning = FALSE}
avg_steps1 <- aggregate(steps~interval + Weekdays, whole_data1, mean)
ggplot(avg_steps1, aes(interval, steps)) +
    geom_line() +
    facet_wrap( ~ Weekdays, nrow = 2) +
    xlab("Interval") +
    ylab("Average Steps")
```

